Higher-Order unification for free!

### 2024-04-16 paper/main.tex
minor

\noindent
The proof found by the solver back-chains on rule 3 (the only rule
about the $\forall$ quantifier), and then solves the premises with
rules 1 and 2 respectively.
Note that rule 3 features a second-order parameter \coqIn{P} that represents
a function of type \coqIn{A ~$\to$~ Prop} (a 1-argument predicate over \coqIn{A}).
The solver has to infer a value for \coqIn{P} by unifying the conclusion
of rule 3 with the goal, and in particular, it has to solve the unification
problem \coqIn{P x = nfact x 3}. This higher-order problem falls in the so-called 
pattern-fragment and admits a unique
solution $\customlabel{solution:intro}{\rho}$ that assigns
the term \elpiIn{~$\lambda$~x.nfact x 3} to \coqIn{P}.
In the rest of this paper shall refer to the pattern-fragment as
\llambda~\cite{miller92jsc}.

@@

\noindent
The main point in showing these equality tests is to remark how
weaker \Ee is, and to identify the four rules that need special
treatment in the implementation of \Uo.
For brevity, we omit the code of \elpiIn{beta}:
it is sufficient to know that <<\elpiIn{beta F L R}>> computes in \elpiIn{R} the
weak head normal form of <<\elpiIn{app [F|L]}>>.
Note that the symbol \elpiIn{|} separates the head of a list from the tail.
\todo{explain forall2}

@@

\begin{proposition}[Simulation]\label{prop:simulation}
$\forall \foUnifPb, \forall \mathcal{N},$ if $~\foUnifPb \subseteq \llambda$
$$
  \runF{\foUnifPb}{\mathcal{N}}{\rho}
  \Leftrightarrow
  \runH{\foUnifPb}{\mathcal{N}}{\rho}
$$
\end{proposition}

>>>>

\noindent
The proof found by the solver back-chains on rule 3 (the only rule
about the $\forall$ quantifier), and then solves the premises with
rules 1 and 2 respectively.
Note that rule 3 features a second-order parameter \coqIn{P} that represents
a function of type \coqIn{A ~$\to$~ Prop} (a 1-argument predicate over \coqIn{A}).
The solver has to infer a value for \coqIn{P} by unifying the conclusion
of rule 3 with the goal, and in particular, it has to solve the unification
problem \coqIn{P x = nfact x 3}. This higher-order problem falls in the so-called 
pattern fragment and admits a unique
solution $\customlabel{solution:intro}{\rho}$ that assigns
the term \elpiIn{~$\lambda$~x.nfact x 3} to \coqIn{P}.
In the rest of this paper shall refer to the pattern fragment as
\llambda~\cite{miller92jsc}.

@@

\noindent
Note that the symbol \elpiIn{|} separates the head of a list from the tail,
that \elpiIn{forall2} applies a binary predicate to two lists (of the same
length). Finally \elpiIn{=>} is $\lambda$Prolog syntax to augment the
current program with an extra rule, and both procedures use it to
assume that the fresh nominal constants introduces by \elpiIn{pi} are
equal to themselves.

The main point in showing these equality tests is to remark how
weaker \Ee is, and to identify the four rules that need special
treatment in the implementation of \Uo.
For brevity, we omit the code of \elpiIn{beta}:
it is sufficient to know that <<\elpiIn{beta F L R}>> computes in \elpiIn{R} the
weak head normal form of <<\elpiIn{app [F|L]}>>.

@@

\begin{proposition}[Run equivalence]\label{prop:simulation}
$\forall \foUnifPb, \forall \mathcal{N},$ if $~\foUnifPb \subseteq \llambda$
$$
  \runF{\foUnifPb}{\mathcal{N}}{\rho}
  \Leftrightarrow
  \runH{\foUnifPb}{\mathcal{N}}{\rho}
$$
\end{proposition}
